---
title: "My Spotify Listening"
output: 
  flexdashboard::flex_dashboard:
    orientation: rows
    source_code: embed
---
<style>                  

  .navbar {
    background-color: #262626;
  }
 .navbar-inverse .navbar-nav>.active>a, .navbar-inverse .navbar-nav>.active>a:hover, .navbar-inverse  .navbar-nav>.active>a:focus >a {
    background-color: #1DB954 !important;
  }
.navbar-inverse .navbar-nav>li>a:hover, .navbar-inverse .navbar-nav>li>a:focus {
    background-color: #464646;
  }
  
</style>
```{r global, include=FALSE}

library(flexdashboard) # r to html
library(jsonlite)      # parse JSON 
library(tidyverse)     # data wrangling
library(highcharter)   # R wrapper for Highcharts
library(DT)            # dynamic tables

setwd("/Users/davidmateos/Downloads/MySpotifyData")
spotify_green <- "#1DB954" 
spotify_dark <- "#464646" 
spotify_gray <- "#6D6D6D" 

## read files delivered by Spotify
listeningHistory1 <- read_json("StreamingHistory0.json", simplifyVector = TRUE)
listeningHistory2 <- read_json("StreamingHistory1.json", simplifyVector = TRUE)
listeningHistory3 <- read_json("StreamingHistory2.json", simplifyVector = TRUE)
listeningHistory  <- as_tibble(do.call("rbind", list(listeningHistory1, 
                                                     listeningHistory2,
                                                     listeningHistory3)))

## Take out what was skipped (only keep 30+ sec)
listeningHistory <- listeningHistory %>% filter(msPlayed > 29999)

## Podcasts I listened to since there are is no is_podcast field
podcasts <- c("The Daily", "Pod Save America", "Lovett or Leave It",
              "The Daily Show With Trevor Noah: Ears Edition", "Setlist",
              "Planet Money", "Keep It!", "The Indicator from Planet Money", 
              "Reply All", "What A Day", "Rough Translation", 
              "The Spurs Show", "Little Gold Men", "Pod Save the World")

```

About
=======================================================================
<br> 

#### <b> Welcome! </b>

<br>
This is a quick, simple dashboard made with R using my Spotify listening activity. 
This data was personally requested and can be similarly <br> retrieved at (https://www.spotify.com/us/account/privacy/) under the 'Donwload Your Data' section. Be warned, this process could take up to a week. 
<br>
<br>

Key things to note about this project:

- Dataset spans from January 10th, 2019 - January 10th, 2020. 

- Key packages used: flexdashboard, highcharter, DT

- The data provided 4 attributes per record: Track Name, Artist Name, Duration of Play, and Time of Play

- Source code is provided above, feel free to replicate with your own data

<br>

Interesting Findings:

- My most listened to genres by artists spent most listening : Pop, Rock, Classic Rock, Latin

- I listen to music the most during the hours of 5-8PM, when I workout

- I listen to podcasts the most at 9am, during my morning commute

- On July 21st, I listened to 20 hours of music!

<br>
<br>
This excercise was inspired by Spotify's 'Year-End Wrapped' campaign and is continuously updated.


<b>Contact:</b> davidumateos@gmail.com


Summary
=======================================================================

Row {data-height=200}
-----------------------------------------------------------------------

### Total

```{r}
total_time <- listeningHistory %>% 
                select(msPlayed) %>% 
                mutate(minPlayed = round(msPlayed/1000/60,3)) %>%
                select(minPlayed) %>% 
                summarise(sum=round(sum(minPlayed)))

total_time <- round(total_time$sum/60) #in hours

valueBox(paste(total_time, "Hours", sep = " "), 
         caption="Listened", 
         icon = "fa-spotify", 
         color = spotify_dark)
```

### Total by Day

```{r}
total_time <- listeningHistory %>% 
                select(msPlayed) %>% 
                mutate(minPlayed = round(msPlayed/1000/60,3)) %>%
                select(minPlayed) %>% 
                summarise(sum=round(sum(minPlayed)))

total_time <- round(total_time$sum/60/365, 1) #in hours

valueBox(paste(total_time, "Hours", sep=" "), 
         caption="Average Daily Listening", 
         icon = "fa-spotify", 
         color = spotify_dark)
```

### Total by Music

```{r}
total_music_time <- listeningHistory %>% 
                      filter(!artistName %in% podcasts) %>%
                      select(msPlayed) %>% 
                      summarise(sum=sum(msPlayed))

total_music_time <- round(total_music_time$sum/1000/60/60) #in hours

valueBox(paste(total_music_time, "Hours", sep = " "),
         caption="Music Played", 
         icon = "fa-music", 
         color = spotify_green)
```

### Total by Podcasts

```{r}
total_podcast_time <- listeningHistory %>% 
                      filter(artistName %in% podcasts) %>%
                      select(msPlayed) %>% 
                      summarise(sum=sum(msPlayed))
total_podcast_time <- round(total_podcast_time$sum/1000/60/60) #in hours

valueBox(paste(total_podcast_time, "Hours", sep = " "), 
         caption="Podcasts Played", 
         icon = "fa-podcast", 
         color = spotify_green)
```

Row {data-height=150}
-----------------------------------------------------------------------

### Top Artist

```{r}

topArtist <- listeningHistory %>% 
                filter(!artistName %in% podcasts) %>%
                 group_by(artistName) %>%
                 summarise(time_listened= round(sum(msPlayed)/1000, 1)) %>%
                 top_n(1) %>% select(artistName)

valueBox(topArtist$artistName, 
         caption="Top Artist", 
         icon = "fa-music", 
         color = spotify_green)
```

### Top Podcast

```{r}

topPod <- listeningHistory %>% 
                filter(artistName %in% podcasts) %>%
                 group_by(artistName) %>%
                 summarise(time_listened= round(sum(msPlayed)/1000, 1)) %>%
                 top_n(1) %>% select(artistName)

valueBox(topPod$artistName, 
         caption="Top Podcast", 
         icon = "fa-podcast", 
         color = spotify_dark)
```

Row {data-height=150}
-----------------------------------------------------------------------

### Top Song 
```{r}
topTrack <- listeningHistory %>% 
                filter(!artistName %in% podcasts) %>%
                 group_by(trackName) %>%
                 summarise(time_listened= round(sum(msPlayed)/1000, 1)) %>%
                 top_n(1) %>% select(trackName)
topTrackArtist <- listeningHistory %>% 
                  filter(trackName == topTrack$trackName) %>% 
                  top_n(1) %>% select(artistName)

topTrack <- gsub("- ", replacement = "",topTrack$trackName)

valueBox(paste(topTrack, topTrackArtist$artistName, sep = " - "),
         caption="Top Track", 
         icon = "fa-music", 
         color = spotify_dark)
```

Row {data-height=150}
-----------------------------------------------------------------------
### Top Episode
```{r}
topEp<- listeningHistory %>% 
                filter(artistName %in% podcasts) %>%
                 group_by(trackName) %>%
                 summarise(time_listened= round(sum(msPlayed)/1000, 1)) %>%
                 top_n(1) %>% select(trackName)
topEpPod <- listeningHistory %>% 
                  filter(trackName == topEp$trackName) %>% 
                  top_n(1) %>% select(artistName)

topEp <- gsub("“" , replacement = "",topEp$trackName)
topEp <- gsub("”", replacement = "",topEp)
valueBox(paste(topEp, topEpPod$artistName, sep = " - "), 
         caption="Top Episode", 
         icon = "fa-podcast", 
         color = spotify_dark)
```

Music
=======================================================================

Row {.tabset .tabset-fade}
-----------------------------------------------------------------------

### Per Day

```{r fig.height= 6}

## take only the date from datetime class and change to EST
musicListeningHistory <- listeningHistory %>% 
                           filter(!artistName %in% podcasts) %>%
                           mutate(datePlayed= 
                                      as.Date(format(as.POSIXct(endTime, tz = "UTC"), "%Y/%m/%d", tz="America/New_York", usetz=T))) %>% 
                           mutate(sPlayed = round(msPlayed/1000/60,2))

musicListeningHistory %>%
  group_by(datePlayed) %>%
  summarise(time= round(sum(sPlayed)/60,1)) %>% 
  hchart(type = "column", hcaes(x = datePlayed, y = time), 
         name = "Time", color = spotify_green) %>%
    hc_xAxis(title = list(text = "")) %>%
    hc_yAxis(title = list(text = "Time Listened (Hrs)")) %>%
    hc_title(text = "When and How much do I listen?", align = "right") %>%
    hc_add_theme(hc_theme_elementary()) 

```

### Per Hour
```{r}

## take only the time from datetime class and change to EST
musicListeningHistory <- listeningHistory %>% 
                            filter(!artistName %in% podcasts) %>%
                            mutate(timePlayed= 
                                     format(as.POSIXct(endTime, tz = "UTC"), "%H", 
                                            tz="America/New_York", 
                                            usetz=F)) %>%
                            mutate(sPlayed = round(msPlayed/1000/60,2))

musicListeningHistory %>%
  group_by(timePlayed) %>%
  summarise(time= round(sum(sPlayed)/60,1)) %>% 
  hchart(type = "column", hcaes(x = timePlayed, y = time), 
         name = "Time", color = spotify_green) %>%
    hc_xAxis(title = list(text = "Time of Day")) %>%
    hc_yAxis(title = list(text = "Time Listened (Hrs)")) %>%
    hc_title(text = "When and How much do I listen?", align = "right") %>%
    hc_add_theme(hc_theme_elementary())
```

Row
-----------------------------------------------------------------------

### Top Artists

```{r}

topArtists <- musicListeningHistory %>%
                 group_by(artistName) %>%
                 summarise(`Time Listened (hr)`= round(sum(sPlayed)/60, 1)) %>%
                 top_n(25) %>% # show the top 25 Artists
                 arrange(-`Time Listened (hr)`) %>% 
                 rename(Artist = artistName)

datatable(topArtists,
          options = list(bPaginate = FALSE,
                         scrollY = FALSE,
                         searching = FALSE,
                         ordering = FALSE,
                         bInfo = FALSE,
                         columnDefs =
                           list(
                             list(
                               className = 'dt-center',
                               targets = c(2))),
                             list(
                               className = 'dt-left',
                               targets = c(1))
                         ),
          class = 'compact')

```

### Top Tracks

```{r}

topTracks <- musicListeningHistory %>%
                 group_by(trackName) %>%
                 summarise(`Time Listened (min)`= round(sum(sPlayed))) %>%
                 top_n(50) %>% # show the top 50
                 arrange(-`Time Listened (min)`) %>% 
                 rename(Track = trackName)

datatable(topTracks,
          options = list(bPaginate = FALSE,
                         scrollY = FALSE,
                         searching = FALSE,
                         ordering = FALSE,
                         bInfo = FALSE,
                         columnDefs =
                           list(
                             list(
                               className = 'dt-center',
                               targets = c(2)),
                             list(
                               className = 'dt-left',
                               targets = c(1)))
                         ),
          class = 'compact')

```

Podcasts
=======================================================================

Row {.tabset .tabset-fade}
-----------------------------------------------------------------------

### Per Day

```{r fig.height= 8}

podcastListeningHistory <- listeningHistory %>% 
                             filter(artistName %in% podcasts) %>%
                             mutate(datePlayed= as.Date(endTime)) %>% 
                             mutate(sPlayed = round(msPlayed/1000/60,2))

podcastListeningHistory %>%
  group_by(datePlayed) %>%
  summarise(time= round(sum(sPlayed)/60,1)) %>% 
  hchart(type = "column", hcaes(x = datePlayed, y = time), 
         name = "Time", color = spotify_green) %>%
    hc_xAxis(title = list(text = "")) %>%
    hc_yAxis(title = list(text = "Time Listened (Hrs)")) %>%
    hc_title(text = "When and How much do I listen?", align = "right") %>%
    hc_add_theme(hc_theme_elementary()) 

```

### Per Hour

```{r}

podcastListeningHistory <- listeningHistory %>% 
                              filter(artistName %in% podcasts) %>%
                              mutate(timePlayed= 
                                         format(as.POSIXct(endTime, tz = "UTC"), "%H", 
                                                tz="America/New_York", 
                                                usetz=F)) %>%
                                mutate(sPlayed = round(msPlayed/1000/60)) 

podcastListeningHistory %>%
  group_by(timePlayed) %>%
  summarise(time= round(sum(sPlayed)/60,1)) %>% 
  hchart(type = "column", hcaes(x = timePlayed, y = time), 
         name = "Time", color = spotify_green) %>%
    hc_xAxis(title = list(text = "Time of Day")) %>%
    hc_yAxis(title = list(text = "Time Listened (Hrs)")) %>%
    hc_title(text = "When and How much do I listen?", align = "right") %>%
    hc_add_theme(hc_theme_elementary())
```

Row
-----------------------------------------------------------------------

### Top Podcasts

```{r}

## add the total time listened by podcast
topPodcasts <- podcastListeningHistory %>%
                 group_by(artistName) %>%
                 summarise(`Total Time Listened (hr)`= round(sum(msPlayed)/1000/3600, 1)) %>%
                 top_n(10) %>% # take the top 10
                 arrange(-`Total Time Listened (hr)`) %>%
                 rename(Podcast = artistName)

## count how many unique episodes were listened to 
## to control for start and stopping an episode group by twice
topPodcastsEpisodes <- podcastListeningHistory %>%
                            group_by(artistName, trackName) %>%
                            summarise(n = n()) %>%
                            group_by(artistName) %>%
                            summarise(n = n()) %>%
                            arrange(-n) %>%
                            rename(Podcast = artistName) %>%
                            rename(Episodes = n)


## join the tibbles above by podcast name
topPodcasts <- topPodcasts %>% 
                  inner_join(topPodcastsEpisodes) %>%
                  mutate(`Avg Episode Length (min)` = 
                           round(`Total Time Listened (hr)`*60/Episodes))

datatable(topPodcasts,
          options = list(bPaginate = FALSE,
                         scrollY = FALSE,
                         searching = FALSE,
                         bInfo = FALSE,
                         columnDefs =
                           list(
                             list(
                               className = 'dt-center',
                               targets = c(2,3,4)),
                             list(
                               className = 'dt-left',
                               targets = c(1)))
                         ),
          class = 'compact')

```
